---
title: "waterr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{waterr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = F, include = F}
library(waterr)
library(chromaticity)
library(ggplot2)
library(dplyr)
```

In this vignette we document the use cases for the various functions within the `waterr` package.

### convert_units

`convert_units` converts the units of water consumption levels. The first argument is the numeric value or vector that you want to convert. The second and third arguments are the units to convert from and to, respectively.

```{r}
# convert_units(100, from = "gallons", to = "hcf")
```

### allocate_consumption

`allocate_consumption` distributes water usage across tiers in a tiered rate structure.

### TierConsumption

This function allocates water consumption across tiers. `nt` is the number of tiers in the rate structure, `usage` is a vector of usage values and `width` inputs are the widths of the various tiers. You only need to input `nt-1` widths as the width of the highest tier is infinite. In the case of water budget rates, `width` parameters can be vectors corresponding to indvidual tier widths.

```{r}
TierConsumption(nt = 2, usage = c(1:20), t1width = 10)
```

### WaterBill

`WaterBill` calculates water bills for customer usage. In this vignette we demonstrate the function on a sample data set of simulted water usage. This function has two main uses, (a) calculating water bills under the current rate structure and (b) calculating water bills under the current and proposed rate structures. The dataframe used on this function, `df` needs to have a column titled `usage` for customer usage.

##### Uniform Rate Structures

When calculating water bills for customers of a water agency, the rates either vary by customer type or they are constant. When calling the function, setting `uniform_rates = T` will apply a standard rate structure across all customers. In this case, be sure to include the rate structure in the call of the function. `fxd_i` refers to fixed rates for rate structures `i = {c, p}` and `vol_i` is a vector of volumetric rates for rate structures `i = {c, p}`.

##### Current Rate Structure Only

If you are interested only in calculating bills under one rate structure you can set `proposed_rates = F` when calling the function. Under this option, bills will be calculated only for the current rate structure.

```{r}
# set.seed(8)
# cust <- data.frame(usage = rbeta(500, 5, 2) * rnorm(500, mean = 10, sd = 5)) %>%
#   mutate(usage = ifelse(usage < 0, 0, usage))
# 
# df <- WaterBill(cust, uniform_rates = T, nt_c = 3,wdths_c = c(5, 5), fxd_c = 20, vol_c = c(4, 5, 6), 
#                 proposed_rates = F) 
# 
# str(df)

```

##### Current and Proposed Rates

Setting `proposed_rates = T` will include bill calculations for two bills as well as some comparisons between the bills from the two rate structures. Under this option, the price elasticity of demand is incorporated into bill calculations in the following steps:

1. The total bill is calculated under the current rate structure as `bill_current`
2. The total bill is calculated under the proposed rate structure as `bill_proposed`
3. Average prices are calculated for both rate structures
4. The new usage level `usage_prime` is estimated based on the change in average price
5. The total bill under the proposed rate structure is calculated with the `usage_prime`
6. The difference between bills between Step 1 and Step 5 is calculated.

```{r}
# # Current Rate Structure
# ntiers_current <- 3
# tierwidths_current <- c(5, 6)
# fixedcharge_current <- 25
# volumetriccharges_current <- c(4, 5, 6)
# 
# # Proposed Rate Structure
# ntiers_proposed <- 4
# tierwidths_proposed <- c(2, 3, 4)
# fixedcharge_proposed <- 30
# volumetriccharges_proposed <- c(2, 3, 6, 7)
# 
# # Elasticity parameter
# el <- -0.1
# 
# df <- WaterBill(df = cust, uniform_rates = TRUE, nt_c = ntiers_current, wdths_c = tierwidths_current, 
#                 fxd_c = fixedcharge_current, vol_c = volumetriccharges_current,
#                 proposed_rates = TRUE, nt_p = ntiers_proposed, wdths_p = tierwidths_proposed, 
#                 fxd_p = fixedcharge_proposed,  vol_p = volumetriccharges_proposed) 
# 
# str(df)
```

