---
title: "waterr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{waterr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = F, include = F}
library(waterr)
library(chromaticity)
library(ggplot2)
library(dplyr)
```

In this vignette we document the use cases for the various functions within the `waterr` package.

### convert_units

`convert_units` converts the units of water consumption levels. The first argument is the numeric value or vector that you want to convert. The second and third arguments are the units to convert from and to, respectively. Unit options are "gallons", "hcf", or "af" (acre-feet).

```{r}
convert_units(100, from = "gallons", to = "hcf")
```

### allocate_consumption

`allocate_consumption` distributes water usage across tiers in a tiered rate structure. `df` should be a dataframe that includes a column for usage in addition to tier widths. The usage column needs to be titled either `use` or `usage`.

Most often, this function will be used in the context of calculating water bills with `calculate_bill`. Because we are typically interested in comparing water bills, we usually need to differentiate between the current and proposed rate structures. This is why we have included the `suffix` argument to the function, which refers to the suffix of tier width columns.

In the example below we have current and proposed widths, as referenced by the suffix.

```{r include = F}
set.seed(8)
df_use <- data.frame(account_number = 1:10, meter_size = c("0.75\"", "1\""), use = round(runif(10, 10, 30), 1), class = "sfr")
df_rates <- data.frame(meter_size = c("0.75\"", "1\""), class = "sfr",
                       t1_width_current = c(10, 11),
                       t1_width_proposed = c(10, 11), t2_width_proposed = c(5, 6),
                       fixed_current = c(20, 30), fixed_proposed = c(21, 32),
                       t1_rate_current = 2, t2_rate_current = 3, t1_rate_proposed = 1.9, t2_rate_proposed = 3, t3_rate_proposed = 5)

df <- left_join(df_use, df_rates, by = c("meter_size", "class")) %>% 
  select(-contains("rate")) %>% 
  select(-contains("fixed"))
```


```{r}
glimpse(df)
```

Let's use `allocate_consumption` for the proposed rates. We see that this function creates the appropriate number of new columns for the dataframe and allocates the customer's usage between them. The suffix of the rate structure is retained in the new usage columns.

```{r}
glimpse(
  df %>% allocate_consumption(suffix = "_proposed")
)
```

When accounting for price elasticity effects during the bill calculation process, we define a new column `use_prime`. If you want to use this column and ignore the original usage column, you can set `use.prime = TRUE`.

### calculate_bill

`calculate_bill` calculates water bills for a set of `_current` rates. The inputs are (1) a `dataframe` for usage and (2) a `dataframe` with rate structure information. If rate structure information is included in `df_use` then set the second argument in the function equal to `NULL`. In the example below we calculate the bill for the current rate structure in our example. We see that the function creates new columns for tiered usage, tiered charges, and the total bill.

```{r}
glimpse(
  calculate_bill(df_use, df_rates)
)
```


### compare_bills

`compare_bills` calculates water bills for a set of `_current` rates and for a set of `proposed` rates. The inputs are (1) a `dataframe` for usage and (2) a `dataframe` with rate structure information. If rate structure information is included in `df_use` then set the second argument in the function equal to `NULL`. 

In the example we calculate the bill for the current rate structure in our example. We see that the function creates new columns for tiered usage, tiered charges, and the total bill for both current and proposed rates. Then, new columns for average price under each set of rates is calculated and usage incorporating price elasticity according to the parameter defined with `eleasticity_par` (`use_prime`). Next, a series of columns for tiered usage and charges for `use_prime` are calculated with the proposed rate structure. `perc_change_bill` captures the percentage change in a customer's bill from the current to proposed rates.

```{r}
glimpse(
  compare_bills(df_use, df_rates)
)
```

### TierConsumption

`TierConsumption` is depracated - see `allocate_consumption`. 

### WaterBill

`WaterBill` is depracated - see `calculate_bill`.  
